---
title: "Item Pool Visualization"
author: "Nils Petras"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Item Pool Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(IPV)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## About this package {#about}

The IPV package is a tool to create IPV charts. The original work on IPV, including the chart concepts, can be found in:
[Dantlgraber, M., Stieger, S., & Reips, U. D. (2019). Introducing Item Pool Visualization: A method for investigation of concepts in self-reports and psychometric tests. Methodological Innovations, 12(3), 2059799119884283.](https://doi.org/10.1177/2059799119884283). Please cite this paper when using the package.
My philosophy is that chart creation should be possible on the full spectrum between quick & dirty and highly customized. So feel free to raise an [issue on github](https://github.com/NilsPetras/IPV/issues) if something frustrates you, or you miss a customization option. Do the same if you encounter a bug (so I can fix it) or do not understand the vignette (so I can change it).

### Quick & dirty chart creation {#quick}
Let us imagine for a second, that estimated SEM parameters are already present in the correct format in an object called "DSSEI":
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- item_chart(data = DSSEI)
mychart
```

As you can see, getting a first result is meant to be trivial. All three IPV chart types can be created by specifying `data = ` only. In most practical cases, I **strongly recommend to** specify a `file_name = `. Results viewed within R (or RStudio) may differ from the file output in multiple ways, including quality. Here, display quality is heavily compromised to generate small files.

Although I have put considerable effort in setting sensible defaults (or generating them dynamically, based on the data), it is always possible to optimize the charts. For example, I prefer my charts to have a higher [data ink ratio](https://infovis-wiki.net/wiki/Data-Ink_Ratio) and like to use [color](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) to guide attention. The many graphical parameters may also enable you to create a black and white print optimized version or a version optimized for small display.

## Workflow
Assuming you have some interesting raw data, empirical, simulated, or faked, these are the necessary steps to produce IPV charts:  
0. SEM estimation  
1. Input  
2. Chart creation  
Tip: Call `?IPV` for an overview of the functions of the package.

### 0. SEM estimation
The IPV package does **not** do the SEM estimation for you. But that means you get to use your favorite statistical software (e.g. the R packages [lavaan](http://lavaan.ugent.be/) or [sem](https://socialsciences.mcmaster.ca/jfox/Misc/sem/SEM-paper.pdf)). [This paper](https://doi.org/10.1177/2059799119884283) shows how to specify the SEM models. For the example, used here and in the paper, you might want to take a look at figures 6-8, which shows the three SEMs used. As the authors point out, the models can only be estimated as SEMs, not by means of factor analysis or in any other way.   
As *nested* I will describe all cases, where the overall item pool is subdivided more than once. For example, an item pool comprising multiple tests can be divided into the tests and subdivided into the tests' facets (as seen in figure 6-8 of the paper). In this case, *nested charts* can be used to compare all three models. In nested cases, I will distinguish a *global* level, where tests are compared with each other, and a nested level, that is concerned with the internal structure of tests. Every case that is not nested, i.e. only consists of a single test, I will call *simple*.

### 1. Input
To use the chart functions, the SEM estimation results need to be formatted in a specific way. This is an example for the data format within the IPV package:
```{r}
self_confidence$tests$RSES
```
Note, that `factor` refers to an item pool, that was divided into subpools (`subfactor`s). In this case, `factor` refers to the Rosenberg Self-Esteem Scale, a test with two facets: Positive Self-Esteem (Ps), and Negative Self-Esteem (Ns), which was reversed here. As seen below, the same data structure applies on the global level, with `factor` referring to the overall self-confidence item pool, comprising the three tests (`subfactor`s) RSES, SMTQ, and DSSEI.  
```{r}
self_confidence$global
```
`cd` is short for center distance  
\begin{equation}
  cd_i = \frac{\lambda^2_{is}}{\lambda^2_{ig}} -1
\end{equation}  
while `mean_cd` is the mean center distance of the items of a facet or test. Furthermore, the matrix of latent correlations between `subfactor`s is given as a second item of the list. To see how the data is combined for nested cases, load the package and call the example object `self_confidence`.  
To spare you the task of creating this data structure by hand, I implemented two automated input pathways. You can either use excel files or the manual input function. In both cases, center distances are calculated automatically and the data is automatically checked for (obvious) errors. Negative center distances are always set to zero before mean center distances are calculated. Regardless of the input mode, you will need to provide:  

* the names of all latent variables and items  
* the factor loadings of the SEMs  
* the latent correlations between the factors within the SEMs.

#### Using manual input functions
These functions allow you to reduce the manual work to a minimum. They are especially useful, when your SEM estimates are already in your R environment (e.g. because you read them from a .csv file). The functions `input_manual_nested()` and `input_manual_simple()` allow you to feed in factor loadings, item names, etc. variable by variable. The correct format is then generated automatically. Run `input_manual_process()` on the result, to automatically calculate center distances.  
This is an example, where all values have been put in individually for demonstration:

```{r}
mydata <- input_manual_simple(
test_name = "RSES",
facet_names = c("Ns", "Ps"),
items_per_facet = 5,
item_names = c(2, 5, 6, 8, 9,
               1, 3, 4, 7, 10),
test_loadings = c(.5806, .5907, .6179, .5899, .6559,
                  .6005, .4932, .4476, .5033, .6431),
facet_loadings = c(.6484, .6011, .6988, .6426, .6914,
                   .6422, .5835, .536, .5836, .6791),
correlation_matrix = matrix(data = c(1, .69,
                                     .69, 1),
                            nrow = 2,
                            ncol = 2))
mydata
input_manual_process(mydata)
```
For nested cases, use the function `input_manual_nested()`, and add the individual tests using `input_manual_simple()`. Then you can run `input_manual_processs()` as in the simple case. You can find the (lengthy) example [below](#nested).  
If any factor loading is below .1 or any center distance below 0, it is set to that value and a warning (or message) is displayed. IPV does not allow negative factor loadings, which is indicated by an error. In this case, recode your data appropriately.

#### Using Excel files
Excel files have the advantage that you can simply copy and paste your SEM estimates into the spreadsheets and the input function of the IPV package (`input_excel()`) does the rest. The files need to be structured as in the example, that you can find here:
```{r}
system.file("extdata", "IPV_global.xlsx", package = "IPV", mustWork = TRUE)
system.file("extdata", "IPV_DSSEI.xlsx", package = "IPV", mustWork = TRUE)
system.file("extdata", "IPV_SMTQ.xlsx", package = "IPV", mustWork = TRUE)
system.file("extdata", "IPV_RSES.xlsx", package = "IPV", mustWork = TRUE)
```

As you can see, there is a file for each test, and a global file. You might want to use a copy as your template, so you can just fill in your values. Open a file to see how it works.  

On sheet 1 you need to provide the factor loadings from your SEM estimation results, on sheet 2 you need to provide the named and complete latent correlation matrix. On sheet 1, "factor" contains a single factor name and "factor_loading" the factor loadings of items on that factor (not squared). "subfactor" contains the names of grouped factors and "subfactor_loading" the factor loadings of items on these factors (not squared). "item" contains the item names. Therefore, each row contains the full information on the respective item.  
Read these excel sheets using input_excel. In the example:
```{r}
global <- system.file("extdata", "IPV_global.xlsx", package = "IPV", mustWork = TRUE)
tests <- c(system.file("extdata", "IPV_DSSEI.xlsx", package = "IPV", mustWork = TRUE),
           system.file("extdata", "IPV_SMTQ.xlsx", package = "IPV", mustWork = TRUE),
           system.file("extdata", "IPV_RSES.xlsx", package = "IPV", mustWork = TRUE))
mydata <- input_excel(global = global, tests = tests)
```

The data will be prepared automatically, including the calculation of center distances. If any factor loading is below .1 or any center distance below 0, it is set to that value and a warning or message is displayed. IPV does not allow negative factor loadings, which is indicated by an error. In this case, recode your data appropriately.

### Facetless tests
In nested charts, tests do not need to have facets. If you use input by excel, use `NA` instead of providing a file name.

```{r, eval=FALSE}
global <- system.file("extdata", "IPV_global.xlsx", package = "IPV", mustWork = TRUE)
tests <- c(system.file("extdata", "IPV_DSSEI.xlsx", package = "IPV", mustWork = TRUE),
           system.file("extdata", "IPV_SMTQ.xlsx", package = "IPV", mustWork = TRUE),
           NA)
mydata <- input_excel(global = global, tests = tests)
```

If you use manual input, do not provide data on facetless tests with `input_manual_simple()`. Any further treatment of facetless tests is handled automatically.


### 2. Chart creation
As briefly outlined above, creating any chart from the formatted data is trivial. There are a few general things to consider, after which I will go into some detail on the three chart types.  

Generally, **best results can be achieved using .pdf files**, since they are vector based. .pdf files can be zoomed and scaled indefinitely without loss of quality. Pixelated formats (.png and .jpeg are supported) will lead to lower quality results and are not scaleable (as can be seen in this vignette). The parameters `file_width = ` and `file_length = ` determine, how large the .pdf file will be, measured in inches (1 in = 2.54 cm). The size of .png or .jpeg files in pixels is determined by multiplying the size in inches with the dots per inch parameter value (`dpi = `). In this vignette the dpi is 72, resulting in heavily pixelated graphics and small file sizes. To inspect your results within RStudio, always use the zoom pop-out of the Plots window, otherwise charts may be heavily distorted. Furthermore, I strongly recommend inspecting the graphics file itself.

Most graphical parameters are size parameters for single elements of the chart, all named `size_... = `,  `width_... = `, or `length_... = `. Those are pretty straightforward: linear scaling parameters defaulting to 1. That means, .5 will half the size and 2 will double it. For all chart types, there is also a global `size = ` parameter, scaling all elements of the chart at once. Use this parameter first, before you fine tune single elements.

For all chart types, it is possible to rotate the whole chart, using `rotate_radians = ` or `rotate_degrees = `. Also, the font can be changed, using the `font = ` parameter. I recommend using the package [extrafont](https://cran.r-project.org/package=extrafont) for access to more fonts.

As you will see, repeatedly trying changes and inspecting the results is necessary to generate the best possible chart. Nevertheless, the final result will be a single, relatively simple function call. I strongly recommend saving your analysis and chart creation functions in a script, so you can always reproduce the results or make changes. If you use [rmarkdown](https://rmarkdown.rstudio.com/) or [sweave](https://rpubs.com/YaRrr/SweaveIntro) to create manuscripts or reports directly from R, changes can easily be made in the scripts and adopted downstream with a single click. Therefore, do not rely on your saved graphics files. 

In the following, I will go into detail on the individual chart types. Using the example, I will show the important customization options provided in the chart functions:

#### Item charts
[Above](#quick), we already saw a first example of an item chart. However, I stated, that I would like to improve the data ink ratio and use color for visual guidance. To reduce the visibility of structural elements, the `fade_... = ` parameters can be used (0 = "black", 100 = "white").
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- item_chart(data = DSSEI,
                      color = "darkblue", color2 = "darkblue",
                      fade_axes = 70, fade_grid_major = 50, fade_grid_minor = 92)
mychart
```
To further accentuate the data, let us change some sizes.
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- item_chart(data = DSSEI,
                      color = "darkblue", color2 = "darkblue",
                      fade_axes = 70, fade_grid_major = 50, fade_grid_minor = 92,
                      size = 1.3, width_items = 1.5, length_items = 1.5, width_grid = .6, size_tick_label = .6)
mychart
```
As you might have noticed, some bars, representing items, are overlapping. This problem is already attenuated, by cropping every other item bar a bit. An alternative is to use different colors or making the item bars slimmer:
```{r, eval=FALSE, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- item_chart(data = DSSEI,
                      color = "darkblue", color2 = "darkred",
                      fade_axes = 70, fade_grid_major = 50, fade_grid_minor = 92,
                      size = 1.3, length_items = 2.5, width_grid = .6, size_tick_label = .6,
                      length_ratio_items = 1, width_items = .9)
mychart
```
A special feature for item charts is the `dodge = ` parameter, that allows long facet labels to dodge the rest of the chart horizontally:
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", fig.show='hold', dev='png'}
x <- DSSEI
colnames(x$cors)[4] <- "Oachkatzlschwoaf"
rownames(x$cors)[4] <- "Oachkatzlschwoaf"
levels(x$cds$subfactor) <- c("Ab", "Pb", "Ph", "Oachkatzlschwoaf")
x$cds$subfactor[16:20] <- "Oachkatzlschwoaf"
mychart1 <- item_chart(data = x)
mychart2 <- item_chart(data = x, dodge = 7)
mychart1
mychart2
```
This works simultaneously for all labels. Labels at the top and bottom do not move, labels on the right and left move the most.

#### Facet charts
Facet charts, as created by `facet_chart()`, can be optimized similarly to item charts, using `fade = ` and `color = `. Furthermore, there are some specific considerations.
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- facet_chart(data = DSSEI)
mychart
```
As you can see in the output, two parameter values (`subradius = ` , and `tick = ` ) have been generated automatically. These need to fit the data. 

The `subradius = ` parameter is important to optimize the appearance. The radius of the facet circles has no meaning. It should be chosen large enough to make the facet labels and correlations readable. But it should also be small enough to make the center distances (thick lines) dominate the first impression and to avoid overlapping facet circles.

For a simplistic version of the chart, the correlations can be omitted.
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- facet_chart(data = DSSEI,
                      cor_labels = FALSE)
mychart
```

In this case, it might be more visually pleasing, to rotate the test label to the top left, and add some color:

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- facet_chart(data = DSSEI,
                      rotate_test_label_radians = pi, color = "firebrick4")
mychart
```

#### Nested charts
Nested charts are the most complex IPV charts. In addition to what I mentioned earlier, there are four important considerations: the `relative_scaling = ` of the global and the nested level, the addition of `xarrows = ` to display correlation arrows between facets of different tests, the ability to `subrotate_... = ` each test individually, and the `cor_spacing = ` to display correlations between the tests. Due to the complexity one should not rely on default values too much:

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- nested_chart(data = self_confidence)
mychart
```
The `relative_scaling = ` should be large enough to have the center distances on the global level shape the overall impression. But a large value for the `relative_scaling = ` makes the nested facet charts of each test small, which should be readable. Note, that the axis scaling within the nested facet charts is different to the global axis scaling by exactly the factor of `relative_scaling = ` , as can be seen from the axis tick marks (small dotted circles). In this particular case, the `relative_scaling = ` seems sound to me, but the facet circles could be larger, including the font sizes within:
```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- nested_chart(data = self_confidence,
                        subradius = .5, size_facet_labels = 2, size_cor_labels_inner = 1.5)
mychart
```
(As you might note, the dynamic default for `relative_scaling = ` adapted to the changes, because the test circles became larger, due to the changes to `subradius = ` .)

The addition of correlation arrows between facets of different tests is indicated by the IPV authors as sensible, when the correlation between these facets exceed the correlation between the respective tests. In the current example, this would result in three extra arrows, that can be added as follows:

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
sc_arrows <- data.frame(test1 = rep(NA, 3), facet1 = NA,
                       test2 = NA, facet2 = NA,
                       value = NA)
sc_arrows[1, ] <- c("DSSEI", "Ab", "RSES", "Ps", ".67")
sc_arrows[2, ] <- c("DSSEI", "Ab", "SMTQ", "Cs", ".81")
sc_arrows[3, ] <- c("SMTQ", "Ct", "RSES", "Ns", ".76")
sc_arrows

mychart <- nested_chart(data = self_confidence,
                        subradius = .5, size_facet_labels = 2, size_cor_labels_inner = 1.5,
                        xarrows = sc_arrows, show_xarrows = TRUE)
mychart
```
The data frame that indicates the names of the facets to connect and the correlation values (here called `sc_arrows` ) needs to be set up with the column names as in the example.

Now the arrows create a lot of overlap and make the chart look messy. This problem can be solved by rotating each of the nested facet charts, so the facets connected by arrows are oriented towards the center. Also the construct label should be moved out of harms way, as well as the test label of the SMTQ.

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- nested_chart(data = self_confidence,
                        subradius = .5, size_facet_labels = 2, size_cor_labels_inner = 1.5,
                        xarrows = sc_arrows, show_xarrows = TRUE,
                        subrotate_degrees = c(180, 270, 90), dist_construct_label = .7,
                        rotate_test_labels_degrees = c(0, 120, 0))
mychart
```

The `cor_spacing = ` refers to the ring around the nested facet charts for each test, in which the correlations between the tests are displayed. It should be large enough for the correlation labels, but not too large. If the correlations are omitted, this ring is also omitted:

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- nested_chart(data = self_confidence,
                        subradius = .5, size_facet_labels = 2, size_cor_labels_inner = 1.5,
                        xarrows = sc_arrows, show_xarrows = TRUE,
                        subrotate_degrees = c(180, 270, 90), dist_construct_label = .7,
                        rotate_test_labels_degrees = c(0, 120, 0),
                        cor_labels_tests = FALSE)
mychart
```
To get a somewhat decent result, let us change some `size_... = ` parameters and add some color. Color can be chosen for the global and the nested level independently. Furthermore, it might be better to increase the line thickness, so the impression of the colored shapes intensifies.

```{r, fig.width=10, fig.height=10, dpi=72, out.height="685px", out.width="685px", dev='png'}
mychart <- nested_chart(data = self_confidence,
                        subradius = .5, size_facet_labels = 2, size_cor_labels_inner = 1.5,
                        xarrows = sc_arrows, show_xarrows = TRUE,
                        subrotate_degrees = c(180, 270, 90), dist_construct_label = .7,
                        rotate_construct_label_degrees = -15,
                        rotate_test_labels_degrees = c(0, 120, 0),
                        color_global = "cyan4", color_nested = "darkblue",
                        size_construct_label = 1.3, size_test_labels = 1.2,
                        width_circles_inner = 1.5, width_circles = 1.5, width_axes_inner = 1.5, width_axes = 1.5)
mychart
```

### Appendix
#### manual input in nested cases - example {#nested}
```{r}
# first the global level
mydata <- input_manual_nested(
  construct_name = "Self-Confidence",
  test_names = c("DSSEI", "SMTQ", "RSES"),
  items_per_test = c(20, 14, 10),
  item_names = c(
     1,  5,  9, 13, 17, # DSSEI
     3,  7, 11, 15, 19, # DSSEI
    16,  4, 12,  8, 20, # DSSEI
     2,  6, 10, 14, 18, # DSSEI
    11, 13, 14,  1,  5,  6, # SMTQ
     3, 10, 12,  8, # SMTQ
     7,  2,  4,  9, # SMTQ
     1,  3,  4,  7, 10, # RSES
     2,  5,  6,  8,  9), # RSES
  construct_loadings = c(
    .5189, .6055, .618 , .4074, .4442,
    .5203, .2479, .529 , .554 , .5144,
    .3958, .5671, .5559, .4591, .4927,
    .3713, .5941, .4903, .5998, .6616,
    .4182, .2504, .4094, .3977, .5177, .4603,
    .3271, .261 , .3614, .4226,
    .2076, .3375, .5509, .3495,
    .5482, .4627, .4185, .4185, .5319,
    .4548, .4773, .4604, .4657, .4986),
  test_loadings = c(
    .5694, .6794, .6615, .4142, .4584, # DSSEI
    .5554, .2165, .5675, .5649, .4752, # DSSEI
    .443 , .6517, .6421, .545 , .5266, # DSSEI
    .302 , .6067, .5178, .5878, .6572, # DSSEI
    .4486, .3282, .4738, .4567, .5986, .5416, # SMTQ
    .3602, .2955, .3648, .4814, # SMTQ
    .2593, .4053, .61  , .4121, # SMTQ
    .6005, .4932, .4476, .5033, .6431, # RSES
    .5806, .5907, .6179, .5899, .6559), # RSES
  correlation_matrix = matrix(data = c(  1, .73, .62,
                                         .73,   1, .75,
                                         .62, .75,   1),
                              nrow = 3,
                              ncol = 3))

# then add tests individually
# test 1
mydata$tests$RSES <- input_manual_simple(
  test_name = "RSES",
  facet_names = c("Ns", "Ps"),
  items_per_facet = c(5, 5),
  item_names = c(2, 5, 6, 8,  9,
                 1, 3, 4, 7, 10),
  test_loadings = c(.5806, .5907, .6179, .5899, .6559,
                    .6005, .4932, .4476, .5033, .6431),
  facet_loadings = c(.6484, .6011, .6988, .6426, .6914,
                     .6422, .5835, .536, .5836, .6791),
  correlation_matrix = matrix(data = c(1, .69,
                                       .69, 1),
                              nrow = 2,
                              ncol = 2))
# test 2
mydata$tests$DSSEI <- input_manual_simple(
  test_name = "DSSEI",
  facet_names = c("Ab", "Pb", "Ph", "So"),
  items_per_facet = 5,
  item_names = c(2, 6, 10, 14, 18,
                  16, 4, 12, 8, 20,
                  3, 7, 11, 15, 19,
                  1, 5, 9, 13, 17),
  test_loadings = c(.302 , .6067, .5178, .5878, .6572,
                    .443 , .6517, .6421, .545 , .5266,
                    .5554, .2165, .5675, .5649, .4752,
                    .5694, .6794, .6615, .4142, .4584),
  facet_loadings = c(.3347, .6537, .6078, .684 , .735 ,
                     .6861, .8746, .7982, .7521, .6794,
                     .7947, .3737, .819 , .7099, .5785,
                     .7293, .8284, .7892, .3101, .4384),
  correlation_matrix = matrix(data = c(1, .49, .66,	.76,
                                       .49,	1, .37, .54,
                                       .66,	.37, 1,	.53,
                                       .76,	.54, .53,	1),
                              nrow = 4,
                              ncol = 4))
# test 3
mydata$tests$SMTQ <- input_manual_simple(
  test_name = "SMTQ",
  facet_names = c("Cf", "Cs", "Ct"),
  items_per_facet = c(6, 4, 4),
  item_names = c(11, 13, 14, 1, 5, 6,
                 3, 10, 12, 8,
                 7, 2, 4, 9),
  test_loadings = c(.4486, .3282, .4738, .4567, .5986, .5416,
                    .3602, .2955, .3648, .4814,
                    .2593, .4053, .61  , .4121),
  facet_loadings = c(.4995, .3843, .5399, .4562, .6174, .6265,
                     .4601, .3766, .4744, .5255,
                     .3546, .5038, .7429, .4342),
  correlation_matrix = matrix(data = c(1,	.71, .62,
                                       .71,	1, .59,
                                       .62,	.59,	1),
                              nrow = 3,
                              ncol = 3))

# finally process (as in a simple case)
my_processed_data <- input_manual_process(mydata)
my_processed_data
```
